An Early Eocene temperature gradient modelled with geological and ecological proxies

### 2022-10-09 manuscript/manuscript.Rmd
lat band segment areal weight solution

## Parameter estimation
We estimated the posterior distributions of the model parameters using a custom-made Markov chain Monte Carlo (MCMC) algorithm, written in R. Specifically, we sampled the unknown parameters $A$, $K$, $M$ and $B$ using Metropolis-Hastings, and used Gibbs sampling to estimate all other unknown parameters. Posterior inference on the modern gradient is based on four chains with 60,000 iterations each, 10,000 of which were discarded as burn-in, and keeping every 10th iteration, resulting in a total of 20,000 iterations. The modern temperature gradient with the Eocene sampling distribution was estimated in one chain with 25,000 iterations for each of the 100 temperature samples. 5,000 iterations each were discarded as burn-in, and every 10th iteration was kept, resulting in a total of 200,000 iterations across all 100 model runs.
For the Eocene model, we ran four chains with 600,000 iterations each, discarding 100,000 as burn-in and keeping every 100th iteration, as the hierarchical model structure results in higher autocorrelation of the chains. The Eocene posterior inference is thus based on a total of 20,000 iterations with low autocorrelation (effective multivariate sample size for $A$, $K$, $M$ and $B$ is > 18,000).

@@

The global Eocene temperatures

>>>>

## Parameter estimation
We estimated the posterior distributions of the model parameters using a Markov chain Monte Carlo (MCMC) algorithm, written in R. Specifically, we sampled the unknown parameters $A$, $K$, $M$ and $B$ with Metropolis-Hastings, and used Gibbs sampling to estimate all other unknown parameters. Posterior inference on the modern gradient is based on four chains with 60,000 iterations each, 10,000 of which were discarded as burn-in, and keeping every 10th iteration, resulting in a total of 20,000 iterations. The modern temperature gradient with the Eocene sampling distribution was estimated in one chain with 25,000 iterations for each of the 100 temperature samples. 5,000 iterations each were discarded as burn-in, and every 10th iteration was kept, resulting in a total of 200,000 iterations across all 100 model runs.
For the Eocene model, we ran four chains with 600,000 iterations each, discarding 100,000 as burn-in and keeping every 100th iteration, as the hierarchical model structure results in higher autocorrelation of the chains. The Eocene posterior inference is thus based on a total of 20,000 iterations with low autocorrelation (effective multivariate sample size for $A$, $K$, $M$ and $B$ is > 18,000).

@@

The global Eocene mean sea surface temperature is estimated at 28.8 (95% CI: 25.7 - 31.7) degC, significantly higher than the modern (?? degC).
